---
title: "Publishing a glasstabs Workflow to Posit Connect"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Publishing a glasstabs Workflow to Posit Connect}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Overview

`runGlassExample("connect-workflow")` opens a Shiny page designed to be
published to Posit Connect. It demonstrates a practical review workflow:

- filters define the active queue with `glassSelect()` and
  `glassMultiSelect()`
- `glassTabsUI(orientation = "vertical", indicator = "solid")` guides users
  through intake, exploration, and approval
- `updateGlassTabBadge()` keeps tab badges aligned with the filtered data
- `updateGlassTabsUI()` moves reviewers to the next stage from server actions
- `theme = "auto"` follows Bootstrap 5 / bslib light and dark modes
- `overflow = "scroll"`, `"wrap"`, and `"menu"` can be compared without
  leaving the app
- opt-in swipes, dynamic tab controls, keyboard behavior, and the experimental
  `glassPage()` wrapper can be exercised in one published session

The page includes an open test checklist. It is there to make a review session
easy to repeat, not to turn the example into a wall of instructions.

## What to test after publishing

1. Resize the page and compare the three overflow choices. The page itself
   should never scroll sideways.
2. On a touch screen, swipe over open panel space. Text inputs, buttons,
   tables, and the sideways scroll box should keep their own gestures.
3. Move through the tab bar with arrow keys, Home, and End. Only the active tab
   should be in the normal Tab-key order.
4. Open both glass selects and try arrow keys, Home, End, Enter, Space, and
   Escape.
5. Add and remove the live tab. Repeat this in compact-menu mode and confirm
   that its option appears and disappears too.
6. Hide and show Approve, then disable and enable Explore. A hidden or disabled
   tab should not trap selection or keyboard focus.
7. Update the Test lab badge. Its pulse should be subtle and should not move
   focus.
8. Compare light and dark mode with each indicator, shape, alignment, and
   orientation.
9. Enable reduced motion or high contrast in the operating system and reload
   the app. Content should change immediately and remain easy to distinguish.
10. Refresh the Connect app and repeat a few controls to catch asset caching or
    session-reconnect problems.

## Run locally

```{r}
library(glasstabs)

if (interactive()) {
  runGlassExample("connect-workflow")
}
```

## Manifest

The example folder ships with a `manifest.json` generated by
`rsconnect::writeManifest()`. The app adapts to the installed `glasstabs`
version: it uses the enhanced responsive arguments when available and falls
back to the CRAN-compatible tab API when Connect restores an older package.
For the v0.4.0 test lab, install the current GitHub development version before
regenerating the manifest so Connect restores the same code you tested:

```{r}
pak::pak("PrigasG/glasstabs")
rsconnect::writeManifest(appDir = "inst/examples/connect-workflow")
```

## Deploy to Posit Connect

The example is a self-contained Shiny app folder with `app.R` and
`manifest.json`. In a package checkout, deploy it directly:

```{r}
rsconnect::deployApp(
  appDir = "inst/examples/connect-workflow",
  appName = "glasstabs-connect-workflow",
  appTitle = "glasstabs Connect Workflow"
)
```

If you are deploying from an installed package, copy the example folder first:

```{r}
example_dir <- system.file("examples", "connect-workflow", package = "glasstabs")
file.copy(example_dir, "connect-workflow", recursive = TRUE)

rsconnect::deployApp(
  appDir = "connect-workflow",
  appName = "glasstabs-connect-workflow",
  appTitle = "glasstabs Connect Workflow"
)
```

## Workflow pattern

The page uses a common Connect pattern: make the first screen the actual work
surface, keep controls near the content they affect, and expose state through
badges and compact metrics. That makes the app useful as a scheduled report,
review queue, or operational handoff page without adding a separate landing
screen.
