---
title: "Responsive and Accessible Tabs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Responsive and Accessible Tabs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Tabs often begin with three short labels and grow with the app. This article
keeps that larger version comfortable on a phone, with a keyboard, and under
system accessibility settings.

## A complete responsive app

```{r complete-app}
library(shiny)
library(glasstabs)

ui <- fluidPage(
  useGlassTabs(),
  glassTabsUI(
    "reports",
    glassTabPanel("summary", "Summary", selected = TRUE,
      h3("Summary"), p("A quick view of the current reporting period.")
    ),
    glassTabPanel("activity", "Recent activity",
      h3("Recent activity"), tableOutput("activity_table")
    ),
    glassTabPanel("quality", "Data quality",
      h3("Data quality"), plotOutput("quality_plot")
    ),
    glassTabPanel("settings", "Team settings",
      h3("Team settings"), textInput("team_name", "Team name")
    ),
    overflow = "scroll",
    theme = "auto"
  )
)

server <- function(input, output, session) {
  output$activity_table <- renderTable(head(mtcars))
  output$quality_plot <- renderPlot(plot(mtcars$wt, mtcars$mpg))
}

shinyApp(ui, server)
```

On a narrow screen, the labels remain in a touch-scrollable row. Keyboard
focus follows the selected tab, and a server update brings its destination
back into view.

## Choosing the overflow style

`"scroll"` is a natural fit for most tab bars. `"multiline"` is useful when every
label should be visible at once. In a horizontal layout, `"menu"` keeps the
header compact and uses a native select control. Vertical tabs already have a
dedicated rail, so compact menu mode is not available there.

```{r overflow-choices}
glassTabsUI(
  "wrapped",
  glassTabPanel("one", "First report", selected = TRUE, p("First")),
  glassTabPanel("two", "Second report", p("Second")),
  glassTabPanel("three", "Third report", p("Third")),
  overflow = "multiline"
)

glassTabsUI(
  "menu",
  glassTabPanel("one", "First report", selected = TRUE, p("First")),
  glassTabPanel("two", "Second report", p("Second")),
  overflow = "menu"
)
```

## Touch swipes without surprises

`swipe = TRUE` adds left and right gestures to ordinary panel content. Inputs,
links, buttons, Shiny outputs, HTML widgets, and horizontal scrollers keep
their own touch behavior.

```{r touch}
glassTabsUI(
  "story",
  glassTabPanel("opening", "Opening", selected = TRUE, p("The first part.")),
  glassTabPanel("middle", "Middle", p("The middle part.")),
  glassTabPanel("ending", "Ending", p("The final part.")),
  swipe = TRUE
)
```

Add `data-gt-no-swipe` to any custom region that should always keep its own
gesture:

```{r no-swipe}
tags$div(`data-gt-no-swipe` = "", custom_touch_component)
```

## Motion and contrast

No R setting is needed for reduced motion. The browser preference removes the
transfer trace and badge pulse, shortens CSS transitions, and shows the next
panel without an animation delay.

Windows forced-colors mode uses system colors and visible focus outlines. When
backdrop blur is unsupported, the controls use solid surfaces instead of
losing contrast.

## Experimental fillable page

`glassPage()` keeps a small full-page app pleasantly short:

```{r glass-page}
ui <- glassPage(
  title = "Operations",
  glassTabsUI(
    "main",
    glassTabPanel("overview", "Overview", selected = TRUE, overview_ui),
    glassTabPanel("alerts", "Alerts", alerts_ui),
    overflow = "scroll",
    theme = "auto"
  )
)
```

It is an experimental wrapper around `bslib::page_fillable()`. Plain
`fluidPage()`, `page_fillable()`, and dashboard layouts remain first-class
homes for every glasstabs widget.
