---
title: "Animated Tabs with glasstabs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Animated Tabs with glasstabs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Overview

`glassTabsUI()` provides an animated glass-morphism tab navigation bar for
Shiny. The active tab is tracked by a sliding glass halo with spring easing,
and a luminous transfer trace sweeps across intermediate tabs during navigation.

## Basic usage

Three functions work together:

- `useGlassTabs()` — injects the CSS and JS once, anywhere in the page UI
- `glassTabsUI()` — renders the full tab bar and content area
- `glassTabPanel()` — defines one tab button and its associated content pane

```{r basic}
library(shiny)
library(glasstabs)

ui <- fluidPage(
  useGlassTabs(),
  glassTabsUI("nav",
    glassTabPanel("overview", "Overview", selected = TRUE,
      h3("Overview"),
      p("This pane is shown first.")
    ),
    glassTabPanel("details", "Details",
      h3("Details"),
      p("Switch to this tab using the button above.")
    ),
    glassTabPanel("settings", "Settings",
      h3("Settings"),
      p("A third tab.")
    )
  )
)

server <- function(input, output, session) {}

if (interactive()) shinyApp(ui, server)
```

## Reading the active tab in the server

The active tab value is pushed to Shiny automatically on every click via
`Shiny.setInputValue`. Access it as `input[["<id>-active_tab"]]`:

```{r server-input}
server <- function(input, output, session) {
  observe({
    req(input[["nav-active_tab"]])
    message("Active tab: ", input[["nav-active_tab"]])
  })

  # Or use the convenience wrapper
  active <- glassTabsServer("nav")
  observe(message("Active: ", active()))
}
```

## Keyboard navigation

Only the selected tab sits in the regular Tab-key order. Left and Right move
through a horizontal strip; Up and Down do the same in a vertical rail. Home
and End reach the first and last available tabs. Hidden and disabled tabs are
left out of that path.

Each tab names its panel through `aria-controls`, and each panel points back
with `aria-labelledby`. Those relationships stay in place when tabs are added
or removed from the server.

## Narrow screens and touch

Scroll mode works well for most apps with several tabs:

```{r overflow-scroll}
glassTabsUI(
  "reports",
  glassTabPanel("summary", "Summary", selected = TRUE, summary_ui),
  glassTabPanel("activity", "Recent activity", activity_ui),
  glassTabPanel("quality", "Data quality", quality_ui),
  glassTabPanel("settings", "Team settings", settings_ui),
  overflow = "scroll"
)
```

The strip accepts touch scrolling and brings the selected tab into view. Use
`overflow = "multiline"` when two or more rows suit the layout. In a horizontal
layout, menu mode keeps the same tab content behind a compact native chooser:

```{r overflow-menu}
glassTabsUI(
  "compact_nav",
  glassTabPanel("queue", "Queue", selected = TRUE, queue_ui),
  glassTabPanel("review", "In review", review_ui),
  glassTabPanel("complete", "Complete", complete_ui),
  overflow = "menu"
)
```

Swipe navigation is deliberately opt-in:

```{r swipe}
glassTabsUI(
  "story",
  glassTabPanel("today", "Today", selected = TRUE, today_ui),
  glassTabPanel("week", "This week", week_ui),
  swipe = TRUE
)
```

Swipes that begin on an input, link, button, Shiny output, HTML widget, or
horizontally scrolling element are ignored. A component can opt out directly
with `data-gt-no-swipe` as well.

The widget follows `prefers-reduced-motion` automatically. In that mode the
panel changes immediately and the transfer trace and badge pulse stay quiet.

## Indicator styles

Use `indicator` to choose how the active tab is drawn:

- `indicator = "glass"` keeps the default frosted halo with shimmer and the
  transfer trace.
- `indicator = "solid"` uses a flat sliding pill with no `backdrop-filter`,
  which is lighter on the GPU for plain dashboards.
- `indicator = "underline"` uses a slim sliding bar and removes the tab pill
  background.

```{r indicators}
glassTabsUI("nav",
  glassTabPanel("overview", "Overview", selected = TRUE, p("Overview")),
  glassTabPanel("data", "Data", p("Data")),
  glassTabPanel("settings", "Settings", p("Settings")),
  indicator = "solid"
)
```

The same tabs can be rendered in all three styles without changing their
content or server logic:

```{r indicator-comparison}
make_tabs <- function(id, indicator) {
  glassTabsUI(
    id,
    glassTabPanel("overview", "Overview", selected = TRUE, p("Overview")),
    glassTabPanel("data", "Data", p("Data")),
    indicator = indicator,
    compact = TRUE
  )
}

tagList(
  make_tabs("glass_tabs", "glass"),
  make_tabs("solid_tabs", "solid"),
  make_tabs("underline_tabs", "underline")
)
```

## Vertical orientation

Set `orientation = "vertical"` to stack the tabs in a left-hand rail beside the
content pane. In vertical mode, Up/Down arrow keys move between tabs and
`indicator = "underline"` becomes a side bar adjacent to the content.

```{r vertical}
glassTabsUI("side",
  glassTabPanel("inbox", "Inbox", selected = TRUE, p("Inbox")),
  glassTabPanel("sent", "Sent", p("Sent")),
  glassTabPanel("archive", "Archive", p("Archive")),
  orientation = "vertical",
  indicator = "underline"
)
```

Use `tab_align = "left"`, `"center"`, or `"right"` to place the tab group in
the available navigation area. This works in horizontal and vertical layouts.
Use `text_align` when the labels and icons inside the buttons should have a
different alignment:

```{r tab-align}
glassTabsUI("aligned",
  glassTabPanel("drafts", "Drafts", selected = TRUE, p("Draft queue")),
  glassTabPanel("review", "Review", p("Review queue")),
  glassTabPanel("done", "Done", p("Completed queue")),
  orientation = "vertical",
  tab_align = "right",
  text_align = "left",
  indicator = "solid"
)
```

## Placing a filter widget beside the tabs

Pass any UI element to `extra_ui` to place it to the right of the tab bar.
The most common use is a `glassMultiSelect()` filter:

```{r extra-ui}
choices <- c(Alpha = "alpha", Beta = "beta", Gamma = "gamma")

ui <- fluidPage(
  useGlassTabs(),
  glassTabsUI("nav",
    extra_ui = glassMultiSelect("cat", choices, show_style_switcher = FALSE),
    glassTabPanel("a", "Tab A", selected = TRUE,
      p("Content A"),
      glassFilterTags("cat")   # tag pills synced to the filter
    ),
    glassTabPanel("b", "Tab B",
      p("Content B"),
      glassFilterTags("cat")
    )
  )
)
```

## Theming

### Built-in presets

```{r theme-presets}
# Dark (default)
glassTabsUI("nav", theme = "dark", ...)

# Light — suits white page backgrounds and bs4Dash cards
glassTabsUI("nav", theme = "light", ...)

# Auto - follows Bootstrap 5 / bslib data-bs-theme light/dark mode
glassTabsUI("nav", theme = "auto", ...)
```

### Custom theme with `glass_tab_theme()`

Supply only the values you want to change — everything else falls back to the
dark preset:

```{r theme-custom}
# Change only the halo colour
glassTabsUI("nav",
  theme = glass_tab_theme(
    halo_bg         = "rgba(251,191,36,0.18)",
    halo_border     = "rgba(251,191,36,0.40)",
    tab_active_text = "#fef3c7"
  ),
  glassTabPanel("a", "Tab", selected = TRUE, p("Content"))
)
```

All eight handles available in `glass_tab_theme()`:

| Argument | What it controls |
|---|---|
| `tab_text` | Inactive tab label colour |
| `tab_active_text` | Active tab label colour |
| `halo_bg` | Sliding glass halo fill |
| `halo_border` | Sliding glass halo border |
| `content_bg` | Tab content panel background |
| `content_border` | Tab content panel border |
| `card_bg` | Inner card background |
| `card_text` | Inner card text colour |

## bs4Dash integration

Set `wrap = FALSE` when embedding inside a `bs4Card` or `bs4Box` — the card
body already provides a constrained container. Pair with `theme = "light"` to
match the card's white background:

```{r bs4dash}
library(bs4Dash)
library(glasstabs)

ui <- bs4DashPage(
  header  = bs4DashNavbar(title = "My App"),
  sidebar = bs4DashSidebar(disable = TRUE),
  body    = bs4DashBody(
    useGlassTabs(),
    bs4Card(
      title = "Analysis", width = 12,
      glassTabsUI("dash",
        wrap     = FALSE,
        theme    = "light",
        extra_ui = glassMultiSelect("f", choices, theme = "light",
                                    show_style_switcher = FALSE),
        glassTabPanel("a", "Overview", selected = TRUE, p("Overview content")),
        glassTabPanel("b", "Details",  p("Detail content"))
      )
    )
  )
)
```

## Multiple tab widgets on one page

Each `glassTabsUI()` instance is scoped to its own `id`, so multiple widgets
work fully independently:

```{r multi-instance}
ui <- fluidPage(
  useGlassTabs(),    # only needed once per page
  glassTabsUI("widget1",
    glassTabPanel("a", "One-A", selected = TRUE, p("Widget 1, pane A")),
    glassTabPanel("b", "One-B", p("Widget 1, pane B"))
  ),
  glassTabsUI("widget2",
    glassTabPanel("x", "Two-X", selected = TRUE, p("Widget 2, pane X")),
    glassTabPanel("y", "Two-Y", p("Widget 2, pane Y"))
  )
)

server <- function(input, output, session) {
  observe(message("Widget 1 active: ", input[["widget1-active_tab"]]))
  observe(message("Widget 2 active: ", input[["widget2-active_tab"]]))
}
```
